---
title: "Viz 3/EDA"
author: "Lucia Wang (lw3061)"
date: "2023-10-05"
output: github_document
---

```{r}
library(tidyverse)
```

import data as before, add month variable
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

## initial numeric work
```{r}
weather_df |>
  ggplot(aes(x=prcp)) +
  geom_histogram()
```
pretty skewed distribution, but some really big outliers.
```{r}
weather_df |>
  filter(prcp > 1000)
```
3 obs bigger than 1000.

```{r}
weather_df |>
  filter(tmax >= 20, tmax <= 30) |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point()
```
banded structure - why? recorded in F, converted to C, rounded? Waterhole is recording it differently which is why its distribution looks different

## grouping
```{r}
weather_df |>
  group_by(name, month)
```
seems invisible (doesn't add columns), but shows that its grouped/how many groups there are in the message. (like sorting). it looks the same, but the df knows its grouped

## counting
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(n_obs = n())
```
`summarize()` works over each group - makes new df with only groups and vars you want

```{r}
weather_df |> 
  count(name, month)
```
not a readable df but does similar to summarize with n

```{r}
weather_df |>
  count(name, month) |>
  pivot_wider(
    names_from = name,
    values_from = n
  )
```
a bit more readable! had to purposely "un"tidy it

## general summaries
```{r}
weather_df |>
  group_by(name) |>
  summarize(
    mean_tmax = mean(tmax)
  )
```
missing values! can solve with `drop_na(tmax)`, or within the actual function with `na.rm=`
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm=TRUE),
    median_tmax = median(tmax, na.rm=TRUE),
    sd_tmax = sd(tmax, na.rm=TRUE)
  )
```
and plot!
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm=TRUE)) |>
  ggplot(aes(x=month, y=mean_tmax, color= name)) +
  geom_point() +
  geom_line()
```

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm=TRUE)) |>
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |>
  knitr::kable()
```
`knitr` package, `kable` function formats it for you in the md file into an actual table!


