---
title: "Visualization with ggplot 1"
author: "Lucia Wang"
date: "2023-09-28"
output: github_document
---

(knitr will do your settings for figures)
```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Using NOAA weather data from the internet
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

## Basic scatterplot
first put in dataset, then the aesthetic mappings, then the geometries
```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
```
`geom_point()` puts a dot for each datapoint (scatterplot)

you can also use pipe to make the graphs. same graph - but helpful if you want to add more functions like filter, select, mutate, other data manip steps etc
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point()
```

if you want to save an object/print it, name the plot
```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```
## advanced plots

add color to names and smooth lines thru each group. 

where you define the aesthetic matters because if you put `color=` in a different place it puts that there instead.

`alpha` = opacity
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha=.3) +
  geom_smooth(se=FALSE)
```

now add facets (panels). 
`.` substitutes for col, then put `~` then row vars
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.3) +
  geom_smooth(se=FALSE) +
  facet_grid(. ~ name)
```

lets try a different plot. using precipitation as the size of points
```{r}
weather_df |>
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point(aes(size= prcp),alpha=.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

if assigning a specific color... careful where you put it in the aes

some fun options - `geom_hex()` for density 
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

line plot (connects the dots)
```{r}
weather_df |>
  filter(name == "USW00022534") |>
  ggplot(aes(x=date, y=tmax)) +
  geom_line() 
```


## univariate
same thought process: dataframe, which var maps onto what, what geometry. 
lets make a histogram...
(use `dodge=` to put the groups separate on the same plot - but maybe not best)
```{r}
ggplot(weather_df, aes(x=tmax)) +
  geom_histogram()
```

or a density plot..
`adjust=` is to smooth ?
```{r}
ggplot(weather_df, aes(x=tmax, fill=name)) +
  geom_density(alpha=0.3, adjust=.75)
```

boxplots
```{r}
ggplot(weather_df, aes(x=name, y=tmax)) +
  geom_boxplot()
```

violin plots (density plot + vertical + mirrored)
```{r}
ggplot(weather_df, aes(x=name, y=tmax)) +
  geom_violin()
```

ridge plot (like violin + density? helpful if lots of categories and shape of distribution matters)
```{r}
ggplot(weather_df, aes(x=tmax, y=name)) +
  geom_density_ridges(scale=.85)
```

## save a plot
make a folder in your directory then use `ggsave()`
```{r}
ggweather = 
  ggplot(weather_df, aes(x=tmax, y=name)) +
  geom_density_ridges(scale=.85)

ggsave("results/ggweather.pdf", ggweather)

```


