---
title: "Visualization with ggplot 1"
author: "Lucia Wang"
date: "2023-09-28"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

## Using NOAA weather data from the internet
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

## Basic scatterplot
first put in dataset, then the aesthetic mappings, then the geometries
```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
```
`geom_point()` puts a dot for each datapoint (scatterplot)

you can also use pipe to make the graphs. same graph - but helpful if you want to add more functions like filter, select, mutate, other data manip steps etc
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point()
```

if you want to save an object/print it, name the plot
```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```
## advanced plots

