---
title: "Viz_2"
author: "Lucia Wang (lw3061)"
date: "2023-10-03"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
```

import data as before
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

start plotting
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.5) +
  labs(
    title = "Temperature plot",
    x= "Min daily temp (degs C)",
    y= "Max daily temp (degs C)",
    color = "Location",
    caption = "Max vs min daily temp in 3 locations; data from RNOAA"
  )
```
use `labs` for labels. can label a variety of different aesthetics, can even name the legends or give titles/captions.

various scales
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.5) +
  labs(
    title = "Temperature plot",
    x= "Min daily temp (degs C)",
    y= "Max daily temp (degs C)",
    color = "Location",
    caption = "Max vs min daily temp in 3 locations; data from RNOAA"
  ) +
  scale_x_continuous(
    breaks = c(-15,0,15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right",
    trans = "sqrt",
    # limits = c(0,10)
  )
```
change stuff on the axes with `scale_x_continuous` and `scale_y_continuous`. can also move axes around with `position`, or transform them with `trans`. you can `filter` data before the plot or during the plot with `limits`.

what about the colors?
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha=.5) +
  labs(
    title = "Temperature plot",
    x= "Min daily temp (degs C)",
    y= "Max daily temp (degs C)",
    color = "Location",
    caption = "Max vs min daily temp in 3 locations; data from RNOAA"
  ) +
  viridis:: scale_color_viridis(discrete= TRUE)
```
`hue` for a range of colors...
or `viridis` palette - works for most people !